### 2020-06-04 writeup/setup.tex
sec 3

\textbf{Problem setup.}
Suppose we have $t$ datasets, where $t$ is a fixed value that does not grow with the feature dimension $p$.
In the high-dimensional linear regression setting (e.g. \cite{HMRT19,BLLT20}), the features of the $k$-th task, denoted by $X_k\in\real^{n_i\times p}$, consists of $n_k$ feature vectors given by $x_1, x_2, \dots, x_{n_k}$.
And each feature $x_i = \Sigma^{1/2}z_i$, where $z_i\in\real^p$ consists of i.i.d. entries with mean zero and unit variance.
The sample size $n_k = \rho_k\times p$ for a fixed value $\rho_k$.
The labels $Y_k = X_k \beta_k + \varepsilon_k$, where $\beta_k$ denotes the linear model parameters and $\varepsilon_k$ denotes i.i.d. noise with mean zero and variance $\sigma^2$.
%Recall that we have $t$ labeled training datasets, denoted by $(X_1, Y_1), (X_2, Y_2), \dots, (X_t, Y_t)$, where $X_i\in\real^{n_i\times p}$ and $Y_i\in\real^{n_i}$ for $1\le i\le t$.
%Following \cite{HMRT19,BLLT20}, we assume that for each task $i = 1,2,\dots,t$,  every feature vector is generated as $x = \Sigma_i^{1/2} z$, where $z\in\real^p$ is a random vector with i.i.d. entries of mean zero and unit variance and $\Sigma_i\in\real^{p\times p}$ is a positive semidefinite matrix.
%Without loss of generality, let the $t$-th task be the target task.

@@

%which can be further decomposed as the bias plus the variance of $\hat{\beta}$.
%In order to relate the btradeoff to properties of the data, we need tight concentration bounds for the bias and variance.
%We consider the high-dimensional regime where $n_i$ is a fixed constant $\rho_i > 1$  multiple of $p$ for every $1\le i\le t$, and $p$ is sufficiently large.
%We focus on a setting where $\rho_t$ is small compared to $\set{\rho_i}_{i=1}^{t-1}$.
%This setting captures the need to add more labeled data to reduce the prediction loss of the target task.
%A well-known result for this setting states that $\te_t(\hat{\beta}_t^{\STL}) = \sigma^2 \cdot \tr[(X_t^{\top}X_t)^{-1}\Sigma_t]$ is concentrated around $\frac {\sigma^2} {\rho_t - 1}$ (e.g. Chapter 6 of \cite{S07}), which scales with the data size and noise level of the target task.
%However, this result only applies to a single task.
%Therefore, our goal is to extend this result to multiple tasks.

@@

%For a matrix $X$, let $\lambda_{\min}(X)$ denote its smallest singular value and $\norm{X}$ denote its spectral norm.

@@

\textbf{Problem statement.}
We focus on predicting a particular task, say the $t$-th task without loss of generality.
Let $\hat{\beta}_t^{\MTL}$ denote the multi-task estimator obtained from the procedure above.
Our goal is to compare the prediction loss of $\hat{\beta}_t^{\MTL}$, defined by
{\small\begin{align*}
		\te(\hat{\beta}_t^{\MTL}) = \exargnob{\set{\varepsilon_i}_{i=1}^t}\exarg{x = \Sigma_t^{1/2} z}{({x}^{\top}\hat{\beta} - {x}^{\top}\beta_t)^2}
		= \exargnob{\set{\varepsilon_i}_i^t}\bignorm{\Sigma_2^{1/2} (\hat{\beta}_t^{\MTL} - \beta_t)}^2,
	\end{align*}}%
%= (\hat{\beta} - \beta_t)^{\top}\Sigma_t(\hat{\beta} - \beta_t)
to the prediction loss $L(\hat{\beta}_t^{\STL})$ of the single-task estimator $\hat{\beta}_t^{\STL} = (X_t^{\top}X_t)^{-1}X_t^{\top}{Y_t}$.
We say there is negative transfer if  $L(\hat{\beta}_t^{\MTL}) > \hat{\beta}_t^{\STL}$, or positive transfer otherwise.

As an example, for the setting of two tasks, we can decompose $L(\hat{\beta}_t^{\MTL})$ into a bias term and a variance term (that scales with noise variance $\sigma_2$) as follows (derived in Appendix \ref{app_proof_sec3}).
%Recall that $\hat{\beta}_t^{\MTL}$ is defined as $BW_t$ after solving equation \eqref{eq_mtl}.
%We can  the test error of $\hat{\beta}_{t}^{\MTL}$ on the target task into two parts as follows.
{\small\begin{align}
	\te(\hat{\beta}_t^{\MTL}) =& ~ \hat{v}^2 \bignorm{\Sigma_2^{1/2} (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_1^{\top}X_1 (\beta_1 - \hat{v}\beta_2)}^2 \label{eq_te_model_shift} \\
	&+ \sigma^2\cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2}, \label{eq_te_var}
\end{align}}%
Here $\hat{v} = W_1 / W_2$ where $W_1, W_2$ is the solutions from solving equation \eqref{eq_mtl_eval} (recalling that $W_1, W_2$ are scalars for two tasks).
The role of $\hat{v}$ is to scale the shared subspace $B$ to fit each task.

Equation \eqref{eq_te_model_shift} corresponds to the bias of $\hat{\beta}_t^{\MTL}$.
Hence, the bias of $\hat{\beta}_t^{\MTL}$ introduces a negative effect that depends on the \textit{similarity} between $\beta_1$ and $\beta_2$.
Equation \eqref{eq_te_var} corresponds to the variance of $\hat{\beta}_t^{\MTL}$, which is always smaller than the variance of $\hat{\beta}_t^{\STL}$.
Intuitively, the more \textit{samples} we add, the smaller the variance is.
%This part introduces a positive variance reduction effect from adding the source labels.
Hence, whether $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$ is determined precisely by the tradeoff between the negative effect of the bias term and the positive effect of the variance term!
%(i) the negative effect from model shift bias.
%(ii) the positive effect from variance reduction;

>>>>

\textbf{Problem setup.}
Suppose we have $t$ datasets, where $t$ is a fixed value that does not grow with the feature dimension $p$.
In the high-dimensional linear regression setting (e.g. \cite{HMRT19,BLLT20}), the features of the $k$-th task, denoted by $X_k\in\real^{n_i\times p}$, consists of $n_k$ feature vectors given by $x_1, x_2, \dots, x_{n_k}$.
And each feature $x_i = \Sigma^{1/2}z_i$, where $z_i\in\real^p$ consists of i.i.d. entries with mean zero and unit variance.
The sample size $n_k $ equals $\rho_k\cdot p$ for a fixed value $\rho_k$.
The labels $Y_k = X_k \beta_k + \varepsilon_k$, where $\beta_k$ denotes the linear model parameters and $\varepsilon_k$ denotes i.i.d. noise with mean zero and variance $\sigma^2$.
%Recall that we have $t$ labeled training datasets, denoted by $(X_1, Y_1), (X_2, Y_2), \dots, (X_t, Y_t)$, where $X_i\in\real^{n_i\times p}$ and $Y_i\in\real^{n_i}$ for $1\le i\le t$.
%Following \cite{HMRT19,BLLT20}, we assume that for each task $i = 1,2,\dots,t$,  every feature vector is generated as $x = \Sigma_i^{1/2} z$, where $z\in\real^p$ is a random vector with i.i.d. entries of mean zero and unit variance and $\Sigma_i\in\real^{p\times p}$ is a positive semidefinite matrix.
%Without loss of generality, let the $t$-th task be the target task.

@@

%which can be further decomposed as the bias plus the variance of $\hat{\beta}$.
%In order to relate the btradeoff to properties of the data, we need tight concentration bounds for the bias and variance.
%We consider the high-dimensional regime where $n_i$ is a fixed constant $\rho_i > 1$  multiple of $p$ for every $1\le i\le t$, and $p$ is sufficiently large.
%We focus on a setting where $\rho_t$ is small compared to $\set{\rho_i}_{i=1}^{t-1}$.
%This setting captures the need to add more labeled data to reduce the prediction loss of the target task.

%However, this result only applies to a single task.
%Therefore, our goal is to extend this result to multiple tasks.

@@

@@

\textbf{Problem statement.}
We focus on predicting a particular task, say the $t$-th task without loss of generality.
Let $\hat{\beta}_t^{\MTL}$ denote the multi-task estimator obtained from the procedure above.
Our goal is to compare the prediction loss of $\hat{\beta}_t^{\MTL}$, defined by
{\small\begin{align*}
		\te(\hat{\beta}_t^{\MTL}) = \exargnob{\set{\varepsilon_i}_{i=1}^t}\exarg{x = \Sigma_t^{1/2} z}{({x}^{\top}\hat{\beta} - {x}^{\top}\beta_t)^2}
		= \exargnob{\set{\varepsilon_i}_i^t}\bignorm{\Sigma_2^{1/2} (\hat{\beta}_t^{\MTL} - \beta_t)}^2,
	\end{align*}}%
%= (\hat{\beta} - \beta_t)^{\top}\Sigma_t(\hat{\beta} - \beta_t)
to the prediction loss $L(\hat{\beta}_t^{\STL})$ of the single-task estimator $\hat{\beta}_t^{\STL} = (X_t^{\top}X_t)^{-1}X_t^{\top}{Y_t}$.
We say there is negative transfer if  $L(\hat{\beta}_t^{\MTL}) > L(\hat{\beta}_t^{\STL})$, or positive transfer otherwise.

As an example, for the setting of two tasks, we can decompose $L(\hat{\beta}_t^{\MTL}) - L(\hat{\beta}_t^{\MTL})$ into a bias term and a variance term (that scales with noise variance $\sigma_2$) as follows (derived in Appendix \ref{app_proof_sec3}).
%Recall that $\hat{\beta}_t^{\MTL}$ is defined as $BW_t$ after solving equation \eqref{eq_mtl}.
%We can  the test error of $\hat{\beta}_{t}^{\MTL}$ on the target task into two parts as follows.
{\small\begin{align}
	\te(\hat{\beta}_t^{\MTL}) - L(\hat{\beta}_t^{\STL}) =& ~ \hat{v}^2 \bignorm{\Sigma_2^{1/2} (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_1^{\top}X_1 (\beta_1 - \hat{v}\beta_2)}^2 \label{eq_te_model_shift} \\
	+&~ \sigma^2 \bigbrace{\bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2} - \bigtr{(X_2^{\top}X_2)^{-1}\Sigma_2}}. \label{eq_te_var}
\end{align}}%
In the above, $\hat{v} = W_1 / W_2$ where $W_1, W_2$ are obtained from solving equation \eqref{eq_mtl_eval} (recalling that $W_1, W_2$ are scalars for two tasks).
The role of $\hat{v}$ is to scale the shared subspace $B$ to fit each task.

Equation \eqref{eq_te_model_shift} corresponds to the bias of $\hat{\beta}_t^{\MTL}$.
Hence, the bias term introduces a negative effect that depends on the \textit{similarity} between $\beta_1$ and $\beta_2$.
Equation \eqref{eq_te_var} corresponds to the variance of $\hat{\beta}_t^{\MTL}$, minus the variance of $\hat{\beta}_t^{\STL}$, which is always negative.
Intuitively, the more \textit{samples} we have, the smaller the variance is.
Meanwhile, \textit{covariate shift} also affects how small the variance can be.
%This part introduces a positive variance reduction effect from adding the source labels.
%Hence, whether $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$ is determined precisely by the tradeoff between the negative effect of the bias term and the positive effect of the variance term!
%(i) the negative effect from model shift bias.
%(ii) the positive effect from variance reduction;
